
```{r, echo=FALSE}
library(readxl)
library(dplyr)
library(ggplot2)
library(stringr)
```

Quantitative Analysis

First we want to analyse whether there is a correlation between the following:

- Gender with the Likert Scale
- Education Level with Likert Scale

```{r}
# load dataset
df <- read.csv("study_data.csv", stringsAsFactors = FALSE)

# get unique values of gender column
unique(df$Gender)

# convert gender to numeric and factor
df$Gender_numeric <- as.numeric(as.factor(df$Gender)) - 1

# check transformation
head(df[, c("Gender", "Gender_numeric")])

# compute correlation between likert scale and gender
cor_result <- cor.test(df$Likert, df$Gender_numeric, method = "pearson")

print(cor_result)

```
The Pearson correlation coefficient of about 0.041 indicates a very weak relationship between Likert scale and the gender of participants.
A t-statistic of 0.357 and a p-value of 0.7222 indicate that the test result is not statistically significant.
This means there is no evidence to suggest that gender is associated with the Likert score in this dataset.

```{r}
#extract unique education values
unique(df$Education)

#factor values
df$Education_factor <- as.factor(df$Education)
#get factor as numeric value
df$Education_numeric <- as.numeric(df$Education_factor)
cor_result <- cor.test(df$Likert, df$Education_numeric, method = "spearman")

print(cor_result)

```
The Spearman's rank correlation coefficient of 0.061 indicates a very weak relationship between Likert scale and education level of the participants.
The p-value of circa 0.595 suggents that there is no statistic significance betweenLikert score and education level.



Correlation between each answer category and likert scale

```{r, echo=FALSE}

#read excel
df_excel <- read_excel("Kategorisierung.xlsx", sheet = 1)
head(df_excel)

#extract likert values
df_excel$Likert <- as.numeric(df_excel$Likert)

#create dummmy variables
dummy_cat <- model.matrix(~ Category - 1, data = df_excel)
dummy_cat_df <- as.data.frame(dummy_cat)

head(dummy_cat_df)

#calc correlation
correlations <- sapply(names(dummy_cat_df), function(cat_col) {
  test <- cor.test(df_excel$Likert, dummy_cat_df[[cat_col]])
  c(correlation = test$estimate, p_value = test$p.value)
})
correlations <- t(correlations)

print(correlations)

```


Descriptive Analysis:

Demographics:

```{r, echo=FALSE}
#remove duplicates
df_unique <- distinct(df, `Participant ID`, .keep_all = TRUE)
nrow(df_unique)

#summarize age from participants
mean_age <- mean(df_unique$Age, na.rm = TRUE)
median_age <- median(df_unique$Age, na.rm = TRUE)
age_summary <- summary(df_unique$Age)

cat("Unique Participant Age Summary:\n")
cat("Mean Age:", mean_age, "\n")
cat("Median Age:", median_age, "\n")
print(age_summary)

#summarize gender frequencies from unique participants
gender_freq <- table(df_unique$Gender)
cat("\nUnique Participant Gender Frequencies:\n")
print(gender_freq)

#summarize education frequencies from unique participants
education_freq <- table(df_unique$Education)
cat("\nUnique Participant Education Frequencies:\n")
print(education_freq)


```
Descriptive Statistics for Likert + Tabulation of Groups

```{r,echo=FALSE}
#Likert
mean_likert <- mean(df$Likert, na.rm = TRUE)
sd_likert   <- sd(df$Likert, na.rm = TRUE)
median_likert <- median(df$Likert, na.rm = TRUE)
summary_likert <- summary(df$Likert)

cat("Descriptive Statistics for Likert Scores:\n")
cat("Mean:", mean_likert, "\n")
cat("Standard Deviation:", sd_likert, "\n")
cat("Median:", median_likert, "\n")
print(summary_likert)

#Groups
group_counts <- table(df$Group)
group_proportions <- prop.table(group_counts)

cat("\nGroup Counts:\n")
print(group_counts)

cat("\nGroup Proportions:\n")
print(group_proportions)

```


Comparative Analysis of Credibility Ratings (Likert)

Now we want to compare the average Likert scores across all groups (Expert/Expert-like/Ordinary Content Creator)

1. ANOVA/Kruskal-Wallis-Test

We have a small sample size, since ANOVA is good for being robust against violatons of normality. (for continuous values)
Kruskal-Wallis (non-parametric) doesn't assume normality for clearly ordinal values.

Since our values are clearly continuous, we will choose ANOVA.

```{r,echo=FALSE}
#factor group variable
df$Group <- as.factor(df$Group)

# run ANOVA to test differences in likert scores by group
anova_result <- aov(Likert ~ Group, data = df)
anova_summary <- summary(anova_result)
print(anova_summary)

# extract the p-value from the ANOVA table 
p_value <- anova_summary[[1]]["Group", "Pr(>F)"]

if(p_value < 0.05) {
  cat("\nANOVA is significant (p =", p_value, "). Performing post-hoc Tukey HSD test:\n")
  posthoc <- TukeyHSD(anova_result)
  print(posthoc)
} else {
  cat("\nANOVA is not significant (p =", p_value, "). No need for post-hoc tests.\n")
}

```
Expert-like Creator vs. Expert:
Difference is -2.87. This means that on average the Likert scale score for Expert-like Creators is 2.87 points lower than for Experts
This difference is statistically significant (p = 0.0005971).

Normal Creator vs. Expert:
Difference is -4.43. This means that on average the Likert scale score for Normal Creators is 4.43 points lower than for Experts.
This is highly significant (p = 0.0000002).

Normal Creator vs. Expert-like Creator:
Difference is -1.56. This means that on average the Likert scale score for Normal Creators is 1.56 points lower than for Expert-like Creators
This is significant (p = 0.0446229).

These results suggest a clear ordering in perceived credibility:

Expert: on avg. highest Likert scale scores

Expert-likes: on avg. second highest Likert scale scores

Ordinary Content Creators: on avg. third highest Likert scale scores



